### 2023-10-27 let.tex
Updates from Overleaf

This is achieved by adding a ``switch type'', denoted
$\TUnknownSwitch$, existing entirely to trigger a ``switch'' to synthesis. $\TUnknownSwitch$ represents a trivial constraint, instantiated as a type for the sole purpose of using the existing machinery to propagate the appropriate mode switch to the relevant sub-expression of the definition. This switch type carries over identically to a non-gradual setting, but here we visually represent as a variant of $\TUnknown$ as it behaves identically with respect to type consistency and matched arrow/product judgements. In the previous example, the unannotated pattern variable $\PVar{a}$ synthesizes
$\TUnknownSwitch$, and the annotated pattern $\PAsc{\PVar{b}}{\TUnknown}$ synthesizes $\TUnknown$.
We write $\ctxSynPatU{\ctx}{\PMV}{\TMV}$ to say that the pattern $\PMV$ synthesizes type $\TMV$.
%
\begin{mathpar}
  \judgment{
    \ctxSynTypeU{\ctx}{\EMV}{\TMV}
  }{
    \ctxAnaTypeU{\ctx}{\EMV}{\TUnknownSwitch}
  }{UASynSwitch}

@@

We omit the marking of patterns---they
may be derived in the same way as those for expressions and are governed by similar metatheorems (again, totality guides the derivation).
In the analytic cases, we introduce error marks that parallel to $\ECInconType{\ECMV}$ and 
$\ECAnaNonMatchedProd{\ECMV}$.

>>>>

This is achieved by adding a ``switch type'', denoted
$\TUnknownSwitch$, existing entirely to trigger a ``switch'' to synthesis. $\TUnknownSwitch$ represents a trivial constraint, instantiated as a type for the sole purpose of using the existing machinery to propagate the appropriate mode switch to the relevant sub-expression of the definition. This switch type carries over identically to a non-gradual setting, but here we represent it as a variant of $\TUnknown$ as it behaves identically with respect to consistency and matched arrow/product judgements. 
%Note that in a non-gradual setting, the matched product rule (or equivalent) would still be required to support the cases such as $\ELet{\PVar{a}}{\EPair{1}{2}}{a}$.
In the previous example, the unannotated pattern variable $\PVar{a}$ synthesizes
$\TUnknownSwitch$, and the annotated pattern $\PAsc{\PVar{b}}{\TUnknown}$ synthesizes $\TUnknown$.
We write $\ctxSynPatU{\ctx}{\PMV}{\TMV}$ to say that the pattern $\PMV$ synthesizes type $\TMV$.
%
\begin{mathpar}
  \judgment{
    \ctxSynTypeU{\ctx}{\EMV}{\TMV}
  }{
    \ctxAnaTypeU{\ctx}{\EMV}{\TUnknownSwitch}
  }{UASynSwitch}

@@

We omit the marking of patterns---they
may be derived in the same way as those for expressions and are governed by similar metatheorems (again, totality guides the derivation).
In the analytic cases, we introduce error marks that parallel to $\ECInconType{\ECMV}$ and 
$\ECAnaNonMatchedProd{\ECMV}$.
\\\\
Note that both let rules assume the let definition synthesizes. In our system this is always the case, but if we added (e.g.) unannotated lambdas, we might wish to relax this restriction. We can freely add variations of the rules where if the definition fails to synthesize, it is instead analyzed against the synthesized type of the pattern. To avoid an additional redundant analysis, we would further want to modify the pattern synthesis judgement to produce a context as well.

### 2023-10-27 related.tex
Updates from Overleaf

% Complex techniques, e.g. based on machine learning or manual weighting, have been proposed to heuristically localize these errors to expressions.

Our focus in this paper has not been on the error messages displayed on screen, about which we make no specific claims. There has been considerable work on improving explanations for type errors, e.g. by providing sample inputs (dynamic witnesses) that elicit runtime errors. With this approach, one can generate graphs for visualizing the execution of witnesses and heuristically identify the source of errors with around 70\% accuracy \cite{Seidel2016}.
These and a variety of other approaches to explaining errors are a fruitful avenue for future integration into the system.

>>>>

Regarding error messages, there has been recent work on using subtyping constraints to model the flow of type information throughout the program to produce more detailed error messages when unification fails \cite{Bhanuka_Parreaux_Binder_Brachthäuser_2023}. Our focus in this paper has not been on the error messages displayed on screen, about which we make no specific claims. However, it may be beneficial in the future to indicate to the user where in the program type suggestions originated from by using type provenances. Documenting the flow of type information would require using asymmetric constraints rather than typical bidirectional equality constraints.

Work on the Helium Haskell compiler \cite{DBLP:conf/haskell/HeerenLI03} explores an approach to constraint-based type inference where different solvers can be selected based on the properties desired of provided feedback. Heeren et al. present two constraint solvers: a global, type graph based constraint solver, which provides detailed error messages and feedback, and a lightweight greedy constraint solver with high performance. When using the global constraint solver, various heuristics can be employed to tune the likelihood of different parts of programs being reported as the source of errors. Additionally, hints can be emitted if corrections are identified that fully resolve type inconsistencies, or if any common sources of errors are identified.

In addition, there has been considerable work on improving explanations for type errors, e.g. by providing sample inputs (dynamic witnesses) that elicit runtime errors. With this approach, one can generate graphs for visualizing the execution of witnesses and heuristically identify the source of errors with around 70\% accuracy \cite{Seidel2016}.
These and a variety of other approaches to explaining errors are a fruitful avenue for future integration into the system.
